---
title: "p8105_hw3_hm2900"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load the data:

```{r}
library(p8105.datasets)
data("instacart")

instacart
```

The dimension of the dataset is `r nrow(instacart)` × `r ncol(instacart)`, which contains 15 variables: `r colnames(instacart)`. 
Some key variables: the variable `r colnames(instacart)[3]` represents the order in which each product was added to cart; the variable `r colnames(instacart)[4]` has the value 1 if this product has been ordered by this user in the past, 0 otherwise; the variable `r colnames(instacart)[6]` represents which evaluation set this order belongs in; the variable `r colnames(instacart)[8]` represents the day of the week on which the order was placed; the variable `r colnames(instacart)[10]` represents days since the last order, capped at 30, NA if order_number=1.


How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
  summarize(
    n_aisle = n_distinct(aisle))

instacart %>%
  group_by(aisle) %>%
  summarize(n_items = n()) %>%
  mutate(items_rank = min_rank(desc(n_items))) %>%
  filter(items_rank == 1)
```

There are 134 aisles. The most items are ordered from the aisle 'fresh vegetables'.  


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r message = FALSE}
instacart %>%
  group_by(aisle) %>%
  summarize(n_items = n()) %>%
  filter(n_items > 10000) %>%
  ggplot(aes(x = n_items, y = aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Number of items",
    y = "Aisle name",
    caption = "Data from p8105 package; The Instacart Online Grocery Shopping Dataset 2017."
  )
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r message = FALSE}
instacart %>%
  group_by(aisle, product_name) %>%
  summarize(n_product = n()) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  mutate(popular_rank = min_rank(desc(n_product))) %>%
  filter(popular_rank %in% c(1, 2, 3)) %>%
  arrange(aisle, popular_rank)
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r message = FALSE}
instacart %>%
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(
    day = case_when(
      order_dow ==  0 ~ "Sunday",
      order_dow ==  1 ~ "Monday",
      order_dow ==  2 ~ "Tuesday",
      order_dow ==  3 ~ "Wednesday",
      order_dow ==  4 ~ "Thursday",
      order_dow ==  5 ~ "Friday",
      order_dow ==  6 ~ "Saturday"
    )
  ) %>%
  select(product_name, day, mean_hour) %>%
  pivot_wider(
    names_from = day,
    values_from = mean_hour
  )
```









