---
title: "p8105_hw3_hm2900"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load the data:

```{r}
library(p8105.datasets)
data("instacart")

instacart
```

The dimension of the dataset is `r nrow(instacart)` × `r ncol(instacart)`, which contains 15 variables: `r colnames(instacart)`. 
Some key variables: the variable `r colnames(instacart)[3]` represents the order in which each product was added to cart; the variable `r colnames(instacart)[4]` has the value 1 if this product has been ordered by this user in the past, 0 otherwise; the variable `r colnames(instacart)[6]` represents which evaluation set this order belongs in; the variable `r colnames(instacart)[8]` represents the day of the week on which the order was placed; the variable `r colnames(instacart)[10]` represents days since the last order, capped at 30, NA if order_number=1.


How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>%
  summarize(
    n_aisle = n_distinct(aisle))

instacart %>%
  group_by(aisle) %>%
  summarize(n_items = n()) %>%
  mutate(items_rank = min_rank(desc(n_items))) %>%
  filter(items_rank == 1)
```

There are 134 aisles. The most items are ordered from the aisle 'fresh vegetables'.  


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r message = FALSE}
instacart %>%
  group_by(aisle) %>%
  summarize(n_items = n()) %>%
  filter(n_items > 10000) %>%
  ggplot(aes(x = n_items, y = aisle, color = aisle)) +
  geom_point() +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Number of items",
    y = "Aisle name",
    caption = "Data from p8105 package; The Instacart Online Grocery Shopping Dataset 2017."
  )
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r message = FALSE}
instacart %>%
  group_by(aisle, product_name) %>%
  summarize(n_product = n()) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  mutate(popular_rank = min_rank(desc(n_product))) %>%
  filter(popular_rank %in% c(1, 2, 3)) %>%
  arrange(aisle, popular_rank)
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r message = FALSE}
instacart %>%
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(
    day = case_when(
      order_dow ==  0 ~ "Sunday",
      order_dow ==  1 ~ "Monday",
      order_dow ==  2 ~ "Tuesday",
      order_dow ==  3 ~ "Wednesday",
      order_dow ==  4 ~ "Thursday",
      order_dow ==  5 ~ "Friday",
      order_dow ==  6 ~ "Saturday"
    )
  ) %>%
  select(product_name, day, mean_hour) %>%
  pivot_wider(
    names_from = day,
    values_from = mean_hour
  )
```

## problem 2

Load the data:

```{r message = FALSE}
accel_data = 
  read_csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(weekend = ifelse(day %in% c("Saturday", "Sunday"), 1, 0)) %>%
  select(week, day_id, day, weekend, everything())

accel_data
```

There are `r ncol(accel_data)` variables and `r nrow(accel_data)` observations in this dataset. Totally there are 5 weeks and each week contains 7 different days with the activity counts for each minute of a 24-hour day starting at midnight. The variable `r colnames(accel_data)[3]` has the value of 0 if the day is a weekday and value of 1 if the day is a weekend. 


Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals.

Are any trends apparent?

```{r message = FALSE}
accel_data %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_minute",
    values_to = "activities"
  ) %>%
  group_by(week, day_id, day, weekend) %>%
  summarize(total_activitiy = sum(activities, na.rm = TRUE)) %>%
  knitr::kable()
```

According to the table, we can see that the total activity number is highest on Sunday in week 1 and on Saturday in week 2, while lowest on Mondays for both of two weeks, which means that for the first two week, there may be more activities on the weekends instead of weekdays. However, for week 3, 4 and 5, the total activity number is highest on Monday, Wednesday, and Friday. Especially for week 4 and 5, the total activity numbers is lowest on Saturday. This indicate that for the further week 3, 4, and 5, there may trend to be more activities on weekdays instead of weekends.


Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.

```{r}
accel_data %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_minute",
    values_to = "activities"
  ) %>%
  separate(activity_minute, into = c("activity", "minute"), sep = 9) %>%
  mutate(minute = as.numeric(minute)) %>%
  ggplot(aes(x = minute, y = activities, color = day)) + 
  geom_line() + 
  labs(
    title = "24-Hour Activity Plot",
    x = "Hour",
    y = "Activities") +
  scale_x_continuous(
    breaks = seq(0, 1440, 60),
    labels = c("12am", "1am", "2am", "3am", "4am", "5am", "6am", "7am", "8am", "9am", "10am", "11am", "12pm", "1pm", "2pm", "3pm", "4pm", "5pm", "6pm", "7pm", "8pm", "9pm", "10pm", "11pm", "12am"))
```

Based on the plot, we can see that the maximum of accelerometer data is between 6am to 11pm over the course of the day, while the minimum is between 12am to 5am. Among the maximum accelerometer data period, there are two peaks existing, one is between 11am to 12 pm on Sunday, and another is between 8pm-10pm on Friday. These may indicate that people prefer to use more accelerometer data before sleeping and when having free time at noon.


